---
title: "Template for contribution to Computo"
subtitle: "Example based on the Rmarkdown system"
author: "the Computo team"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: yeti
    highlight: tango
    toc: true
    toc_depth: 2
bibliography: template-computo-Rmarkdown.bib
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy = FALSE, fig.width = 8, fig.height = 8, echo = TRUE)
options(htmltools.dir.version = FALSE)
```

# Introduction

## About this document

This document provides a Rmarkdown^[https://rmarkdown.rstudio.com/] template for contributions to the **Computo** Journal. We show how `R` or `Python` code can be included.

## Advice in writting your manuscript

First make sure that you are able to build your manuscript as a regular notebook on your system. Then you can start configure the binder environment (which we configure to use an Ubuntu machine with latest LTS release).

# Formating the notebook

This section is about writing a notebook with the Rmarkdown system, typically for R users.

## Rmarkdown basics

We first quickly cover the most basic features of Rmarkdown, that is, formatting text with markdown, math with $\LaTeX$ via MathJax and bibliographical references via $Bib\TeX$. 

[Rmarkdown](http://rmarkdown.rstudio.com) is a simple formatting system for authoring HTML and PDF documents, 
that relies on the markdown markup language.

To render the document as HTML within Rstudio, click the **Knit** button. A document will be generated that includes both content as well as the output of any embedded R code chunks within the document. Alternatively, the shortcut Ctrl + Maj + K will produces the same result.

### Mathematical formulae

$\LaTeX$ code is natively supported, which makes it possible to use mathematical formulae:

$$
f(x_1, \dots, x_n; \mu, \sigma^2) = 
\frac{1}{\sigma \sqrt{2\pi}} \exp{\left(- \frac{1}{2\sigma^2}\sum_{i=1}^n(x_i - \mu)^2\right)}
$$

### References

References are displayed as footnotes using bibtex, e.g. `[@computo]` will display as [@computo], where `computo` is the bibtex key for this entry. The bibliographic information is automatically retrieved from the `.bib` file specified in the header of this document (here: `template-computo-Rmarkdown.bib`).

## R code

`R` code [@R-base] chunks may be embedded as follows:

```{r cars}
knitr::kable(summary(cars))
```

### Including Plots

Plots can be generated as:

```{r pressure, echo=FALSE}
library("ggplot2")
p <- ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth()
p
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

Interative plots may also be produced in the HTML output of the document:

```{r, message=FALSE}
library("plotly")
ggplotly(p)
```

## Python Code

The R package [reticulate](https://cran.r-project.org/package=reticulate) [@R-reticulate] includes a Python engine for R Markdown that enables easy interoperability between Python and R chunks. Below we demonstrate a small subset of the available functionalities. We refer to the vignette [R Markdown Python Engine](https://cran.r-project.org/web/packages/reticulate/vignettes/r_markdown.html) for a more detailed description.

### Setup

```{r load-reticulate}
library("reticulate")
use_virtualenv("computo-template")
```

First make sure (here, in `R`) that the required python modules are available

```{r install-python-modules, eval=TRUE, results="hide"}
if (!py_module_available("seaborn")) py_install("seaborn")
if (!py_module_available("pandas")) py_install("pandas")
if (!py_module_available("matplotlib")) py_install("matplotlib")
```

### Using python

Example of python code and associated output:

```{python iris-python, echo = TRUE}
import pandas as pd
import seaborn as sns
import matplotlib.pyplot as plt

sns.set(style="whitegrid", palette="muted")

# Load the example iris dataset
iris = sns.load_dataset("iris")

# "Melt" the dataset to "long-form" or "tidy" representation
iris = pd.melt(iris, "species", var_name="measurement")
plt.figure()

# Draw a categorical scatterplot to show each observation
sns.swarmplot(x="measurement", y="value", hue="species", palette=["r", "c", "y"], data=iris)
plt.show()
```

## Communication between R and python chunks

All objects created within Python chunks are available to R using the `py` object exported by the reticulate package, e.g.:

```{r iris-r, echo=TRUE}
rmarkdown::paged_table(py$iris)
```

Conversely, all objects created within R are available from Python using the `r` object exported by the reticulate:

First, let us create an object within R:

```{r rvolcano-r, echo=TRUE}
data(volcano)
```

```{r volcano-table}
rmarkdown::paged_table(as.data.frame(volcano))
```

This object is accessible from Python:

```{python volcano-python, echo=TRUE}
print(r.volcano)
```

## Other languages

Theoretically, you can include many others languages into Rmarkdown including Julia and C++. If you are comfortable enough to configure binder and prove us the reproducibility of your code, feel free to use any other language.

# Session information

```{r session-info}
sessionInfo()
```

# References
